\section{Operads and props}

We consider the category $\Ch$ of chain complexes of abelian groups as our base category, remarking that all definitions in this section apply to general closed symmetric monoidal categories.

\subsection{$\S$-modules and $\S$-bimodules}
Recall that a group $G$ can be thought of as a category with a single object and only invertible morphisms. From this viewpoint, a left $G$-module (resp. right $G$-module or $G$-bimodule) is the same as a functor from $G$ (resp. $G^\op$ or $G \times G^\op$) to $\Ch$.

Let $\S$ be the category whose objects are the natural numbers and whose set of morphisms between $m$ and $n$ is empty if $m \neq n$ and is otherwise the symmetric group $\S_n$.
A \textit{left $\S$-module} (resp. right $\S$-module or $\S$-bimodule) is a covariant functor from $\S$ (resp. $\S^\op$ or $\S \times \S^\op$) to $\Ch$.
In this paper we prioritize left module structures over their right counterparts. As usual, taking inverses makes both perspectives equivalent.

The homomorphisms $\S_n \to \S_n \times \S_1$ and $\S_n^\op \to \S_1 \times \S_n^\op$ induce natural forgetful functors $\Uleft$ and $\Uright$ from the category of $\S$-bimodules to those of left and right $\S$-modules.

Given a chain complex $C$ define:
\begin{align*}
\End^C(r) &= \Hom(C, C^{\otimes r})
& \End_C(r) &= \Hom(C^{\otimes r}, C)
&\End^C_C(r, s) &= \Hom(C^{\otimes r}, C^{\otimes s})
\end{align*}
with their natural structures of left $\S$-module, right $\S$-module, and $\S$-bimodule respectively.
The natural forgetful functors from $\S$-bimodules to left and right $\S$-modules send $\End^C_C$ to $\End^C$ and $\End_C$ respectively.

\subsection{Composition structures}

Operads an props are $\S$-modules and \mbox{$\S$-bimodules} respectively enriched with certain composition structures. These are best understood by abstracting the composition structure naturally present in the $\S$-module $\End^C$, naturally an operad, and the $\S$-bimodule $\End^C_C$, naturally a prop.

Succinctly, an operad $\mathcal O$ is an $\S$-module together with a collection of $R$-linear maps
\begin{equation*}
\mathcal O(r) \otimes \mathcal O(s) \to \mathcal O(r+s-1)
\end{equation*}
satisfying suitable associativity, equivariance and unitality conditions.
A prop $\mathcal P$ is an $\S$-bimodule together with two types of compositions; horizontal
\begin{equation*}
\mathcal P(r_1, s_1) \otimes \mathcal P(r_2, s_2) \to \mathcal P(r_1 + r_2, s_1 + s_2)
\end{equation*}
and vertical
\begin{equation*}
\mathcal P(r,s) \otimes \mathcal P(s, t) \to \mathcal P(r, t)
\end{equation*}
satisfying their own versions of associativity, equivariance and unitality.
For a complete presentation of these concepts we refer to Definition 11 and 54 of \cite{Markl08}.

We remark that the compositional structure of a prop $\mathcal P$ restricts to operad structures on $\Uleft(\mathcal P)$ and $\Uright(\mathcal P)$.

\subsection{Representations}

A morphisms of operads or of props is simply a morphisms of their underlying $\S$-modules or $\S$-bimodules preserving the respective compositional structures.

Consider a chain complex $C$, an operad $\mathcal O$ and a prop $\mathcal P$. An $\mathcal O$-\textit{coalgebra} (resp. $\mathcal O$-\textit{algebra}) structure on $C$ is an operad morphism $\mathcal O \to \End^C$ (resp. $\mathcal O \to \End_C$), and a $\mathcal P$-\textit{bialgebra} structure on $C$ is a prop morphism $\mathcal P \to \End_C^C$.

We remark that the linear duality functor naturally transforms an $\mathcal O$-coalgebra structure on a chain complex into an $\mathcal O$-algebra structure on its dual.

\subsection{Free operads and props}
\label{ss:free props}

As described for example in \cite{Markl08}, the \textbf{free prop} $F(M)$ generated by a \mbox{$\S$-bimodule} $M$ is constructed using open directed graphs with no directed loops that are enriched with a labeling described next. We think of each directed edge as built from two compatibly directed half-edges. For each vertex $v$ of a directed graph $G$, we have the sets $in(v)$ and $out(v)$ of half-edges that are respectively incoming to and outgoing from $v$. Half-edges that do not belong to $in(v)$ or $out(v)$ for any $v$ are divided into the disjoint sets $in(G)$ and $out(G)$ of incoming and outgoing external half-edges. For any positive integer $n$, let $\overline{n} = \{1,\dots,n\}$ and $\overline{0} = \emptyset$. For any finite set $S$, denote the cardinality of $S$ by $|S|$. The labeling is given by bijections  
\begin{equation*}
\overline{|in(G)|}\to in(G), \qquad
\overline{|out(G)|}\to out(G),
\end{equation*}
and
\begin{equation*}
\overline{|in(v)|}\to in(v), \qquad
\overline{|out(v)|}\to out(v),
\end{equation*}
for every vertex $v$.
We refer to the isomorphism classes of such labeled directed graphs with no directed loops as $(n,m)$\textbf{-graphs} denoting the set of these by $\G(m,n)$.
We use graphs immersed in the plane to represent elements in $\G(m,n)$, please see Figure \ref{f:immersion}.
We consider the right action of $\S_n$ and the left action of $\S_m$ on a $(n,m)$-graph given respectively by permuting the labels of $in(G)$ and $out(G)$. This action defines the $\S$-bimodule structure on the free prop
\begin{equation} \label{e:free prop}
F(M)(m,n) \ = \bigoplus_{\Gamma \in \G(m,n)} \bigotimes_{v \in Vert(\Gamma)} out(v) \otimes_{\S_q} M(p, q) \otimes_{\S_p} in(v),
\end{equation}
where we simplified the notation writing $p$ and $q$ for $\overline{|in(v)|}$ and $\overline{|out(v)|}$ respectively. The composition structure is defined by (relabeled) grafting and union.

\input{immersion}

We remark that the free operad construction is described analogously using either graphs with a single input or output and only grafting.

\subsection{The prop $\M$}
\label{ss:the prop M}
Let us consider the free prop $F(N)$ generated by the $\S$-bimodule $N$ whose only non-zero chain complexes are concentrated in degree $0$ and are give by
\begin{equation*}
N(1, 0)_0 = \Z\{\varepsilon\}, \qquad
N(1, 2)_0 = \Z[\S_2]\{\Delta\}.
\end{equation*}
Define $\A$ as the quotient of $F(N)$ by the prop ideal generated by the relations
\begin{equation*}
\begin{tikzpicture}[scale=.2]
\draw (-4,0)--(-4,2)--(-5,3)--(-5,4);
\draw (-6,0)--(-6,2)--(-5,3)--(-5,4);
\node [left, scale=.8] at (-5.3,3.2) {$\Delta$};
\draw [fill] (-6,.2) circle [radius=.2];
\node [left] at (-6,0) {$\varepsilon$};

\node at (-2,2) {=};
\draw (0,0)--(0,4);
\node at (2,2) {=};

\draw (4,0)--(4,2)--(5,3)--(5,4);
\draw (6,0)--(6,2)--(5,3)--(5,4);
\node [right, scale=.8] at (5.3,3.2) {$\Delta$};
\draw [fill] (6,.2) circle [radius=.2];
\node [right] at (6,0) {$\varepsilon$.};
\end{tikzpicture}
\end{equation*}
We remark that representations of the prop $\A$ correspond to counital coalgebra.

Let $W^{(1)}$ be the chain complex of free $\Z[\S_2]$-modules
\begin{equation*}
\begin{tikzcd}
\Z[\S_2]\{\nu\} &[0pt] \arrow[l, "1-T"'] \Z[\S_2]\{\mu\},
\end{tikzcd} 
\end{equation*}
which is isomorphic to the cellular chains on the standard model of the circle
\begin{equation*}
\begin{tikzpicture}[scale=.85]
\draw (0,0) circle (20pt);
\node[scale=.7] at (0,12pt){$\mu$};
\node[scale=.7] at (0,-12pt){$T \mu$};
\node[scale=.7] at (-28pt,0){$T \nu$};
\node[scale=.7] at (26pt,0){$\nu$};
\draw [fill] (-20pt,0) circle [radius=1pt];
\draw [fill] (20pt,0) circle [radius=1pt];
\end{tikzpicture}
\end{equation*}
and let $W^{(0)}$ be the subcomplex generated by $\nu$. We think of $W^{(1)}$ as an $\S_2$-equivariant 1-cell with boundary $W^{(0)}$.

We regard these complexes as $\S$-bimodules concentrated in biarity $(2,1)$, and let $\varphi \colon W^{(0)} \to \A$ be define by sending $T \nu$ and $\nu$ respectively to
\begin{equation*}
	\begin{tikzpicture}[scale=.2]
	\draw (-4,0)--(-4,4);
	\draw (-6,0)--(-6,4);
	\draw [fill] (-6,.2) circle [radius=.2];
	\node [left] at (-6,0) {$\varepsilon$};
	
	\node at (0,.4) {and};
	
	\draw (4,0)--(4,4);
	\draw (6,0)--(6,4);
	\draw [fill] (6,.2) circle [radius=.2];
	\node [right] at (6,0) {$\varepsilon$.};
	\end{tikzpicture}
\end{equation*}
Consider the push-out
\begin{equation*}
\begin{tikzcd}
F(W^{(0)}) \arrow[r, "F(\varphi)"] \arrow[d] & \A \arrow[d, dashed] \\
F(W^{(1)}) \arrow[r, dashed] & \mu \vee \A
\end{tikzcd}
\end{equation*}
in the category of props. We think of $\mathcal A \vee_\varphi \mu$ as the prop obtained by attaching a $1$-cell in biarity $(2,1)$ to $\A$.

Define $\M$ as the quotient of $\mu \vee \A$ by the ideal generated by
\begin{equation*}
\begin{tikzpicture}[scale=.2]
\draw (5,4)--(5,3)--(6,2)--(6,0);
\draw (7,4)--(7,3)--(6,2);
\node [left] at (5.5,2) {$\mu$};
\draw [fill] (6,.2) circle [radius=.2];
\node [right] at (6,0) {$\varepsilon$};

\node at (9,2) {= 0.};
\end{tikzpicture}
\end{equation*}

We can give a more explicit description of $\M$ using the constructions of Subsection~\ref{ss:free props}.

An element in $\M(m, n)$ is a linear combination of $(m, n)$-graphs where with vertices of bi-valence $(1,0)$, $(1,2)$ and $(2,1)$ ...

\subsection{The Hopf prop structure} We now describe a diagonal map $\M(m,n) \to \M(m,n) \otimes \M(m,n)$ for every biarity $(m,n)$, compatible with the composition structure of $\M$.