% !TEX root = ../cobar1.tex

\section{Adams's model of based loop spaces} \label{s:theorem1}

%The goal of this section is to prove \cref{t:1st main thm in the intro}.
%Explicitly, we construct a natural $E_{\infty}$-bialgebra structure on $\cobar \schainsA(X)$ for any reduced simplicial set $X$ (\cref{l:lift of cobar to e-infty}); and show that when $X$ is the singular complex of a topological space $(Z,z)$, Adams' comparison map $\theta_Z \colon \cobar \sSchainsA(Z,z) \to \cSchains(\loops_z Z)$ is a quasi-isomorphism of $E_\infty$-bialgebras (\cref{l:adams comparison is an e-infty bialgebra map}).
%
%The first goal is achieved using a categorical reformulation of Baues' geometric cobar construction (\cref{ss:cubical cobar}) together with the natural $\UM$-coalgebra structure defined on cubical chains (\cref{ss:e-infty on cubical}) and its monoidal properties (\cref{l:cubical e-infty chains are monoidal}).
%The second goal is obtained by factoring Adams' comparison map through this cubical model (\cref{ss:factorization of adams}).

\subsection{The cobar construction} \label{ss:cobar construction}

A \textit{coaugmentation} on a coalgebra $C$ is a morphism $\nu \colon \k \to C$ in $\coAlg$.
A coalgebra is said to be \textit{coaugmented} if it is equipped with a coaugmentation.

We denote by $\coAlg^\ast$ the category of coaugmented coalgebras with morphisms being coaugmentation preserving coalgebra morphisms.

The \textit{cobar construction} is the functor
\[
\cobar \colon \coAlg^\ast \to \Mon_{\Ch}
\]
defined on objects as follows.
Let $(C, \partial, \Delta, \varepsilon, \nu)$ be the data of a coaugmented coalgebra.
Denote by $\overline{C}$ the cokernel of $\nu \colon \k \to C$ and recall that $s^{-1}$ is the $(-1)^\th$ suspension.
The cobar construction $\cobar C$ of this coaugmented coalgebra is the graded module
\[
T(s^{-1} \overline{C}) = \k \oplus s^{-1}\overline{C} \oplus (s^{-1}\overline{C})^{\otimes 2} \oplus (s^{-1}\overline{C})^{\otimes 3} \oplus \dots
\]
regarded as a monoid in $\Ch$ with $\mu \colon T(s^{-1} \overline{C})^{\otimes 2} \to T(s^{-1} \overline{C})$ given by concatenation, $\eta \colon \k \to T(s^{-1} \overline{C})$ by the obvious inclusion, and differential constructed by extending the linear map
\[
- s^{-1} \circ \partial \circ s^{+1} \ + \ (s^{-1} \otimes s^{-1}) \circ \Delta \circ s^{+1} \colon
s^{-1} \overline{C} \to s^{-1}\overline{C} \oplus (s^{-1}\overline{C} \otimes s^{-1}\overline{C}) \hookrightarrow T(s^{-1}C)
\]
as a derivation.

In this article we are mostly concerned with \textit{connected} coalgebras, namely coaugmented coalgebras that are non-negatively graded and for which the coaugmentation induces an isomorphism $\k \cong C_0$ of $\k$-modules.
We denote by $\coAlg^0$ the full subcategory of $\coAlg^\ast$ consisting of connected coalgebras, and remark that if $C \in \coAlg^0$, then $ \cobar(C)$ is concentrated on non-negative degrees.

The object $ \cobar(C) \in \Mon_{\Ch}$ may be equipped with a natural augmentation map $\cobar(C) \to \k$ in $\Mon_{\Ch}$ given by the projection map to the first summand term.
Therefore, the cobar construction may be regarded as a functor from coaugmented coalgebras to augmented algebras.
However, we will not require the data of augmentations in this article.

We now review the central examples of coaugmented coalgebras to which Adams applied the cobar construction.

\subsection{Adams' map} \label{ss:adams maps}

In \cite{adams1956cobar}, Adams constructed a natural map of algebras
\begin{equation} \label{e:adams map 2}
	\theta_Z \colon \cobar \sSchainsA(Z,z) \to \cSchains(\loops_z Z),
\end{equation}
for any pointed topological space $(Z,z)$.
The construction is based on a collection of continuous maps
\[
\theta_n \colon \gcube^{n-1} \to P(\gsimplex^n;0,n),
\]
where $P(\gsimplex^n;0,n)$ denotes the topological space of Moore paths in $\gsimplex^n$ from the $0^\th$ vertex $v_0$ to the $n^\th$ vertex $v_n$.
These maps are constructed inductively so that following equations are satisfied:
\begin{enumerate}
	\item $\theta_1(0)\colon \gcube^1 \to \gsimplex^1$ is the path $\theta_1(0)(s) = sv_1 +(1-s)v_0$.
	\item $\theta_n \circ e_j^0 = P(d_j) \circ \theta_{n-1}$.
	\item $\theta_n \circ e_j^1 = (P(f_j) \circ \theta_j) \cdot (P(l_{n-j}) \circ \theta_{n-j})$.
\end{enumerate}
In the above equations $f_j\colon \gsimplex^j \rightarrow \gsimplex^n$ and $l_{n-j}\colon \gsimplex^{n-j} \rightarrow \gsimplex^n$ denote the first $j^\th$ face map and last $(n-j)^\th$ face map of $\gsimplex^n$, respectively, and
$e_j^0,e_j^1\colon \gcube^{n-1} \rightarrow \gcube^{n}$ denote the $j^\th$ bottom and top cubical face maps, respectively.
For any continuous map of spaces $f \colon Z \to Z'$, we denote by $P(f) \colon P(Z) \to P(Z')$ the induced map at the level of spaces of paths.
Given any two composable paths $\alpha$ and $\beta$, the dot symbol in $\alpha \cdot \beta$ denotes the composition (or concatenation) of paths.

Adams' map $\theta_Z$ is now given as follows.
For any singular $1$-simplex $\sigma \in S^{\Delta}_1(Z,z)$ let
\[
\theta_Z([\sigma]) = P(\sigma) \circ \theta_1 - c_z,
\]
where $c_z \in \cSchains_0(\loops_z Z)$ is the singular $0$-cube determined by the constant loop at $z \in Z$.
For any singular $n$-simplex $\sigma \in S^{\Delta}_n(Z,z)$ with $n>1$, let
\[
\theta_Z([\sigma]) = P(\sigma) \circ \theta_n.
\]
Since the underlying graded algebra of $\cobar \sSchainsA(Z,z)$ is free, we may extend the above to an algebra map $\theta_Z \colon \cobar \sSchainsA(Z,z) \to \cSchains(\loops_z Z)$.
The compatibility equations for the maps $\theta_n$ imply that $\theta_Z$ is a chain map.

We will construct an $E_{\infty}$-bialgebra structure on $\cobar \sSchainsA(Z,z)$ and show that, when $\cSchains(\loops_z Z)$ is equipped with the $E_{\infty}$-coalgebra structure described in \cref{ss:e-infty on cubical}, the map $\theta_Z$ is a map of $E_{\infty}$-coalgebras.
With this goal in mind, we describe a categorical version of Adams' constructions.

